A comparison of the most popular AI text-to-image generation applications

### 2023-01-26 Paper.tex
Commit

	\begin{abstract}
		Following the explosion of artificial intelligence during the year of 2022, there are now many text-to-image generation applications out there. This is great for the field of artificial intelligence as a whole, however it has made knowing which application is best quite hard for artists who are new to text-to-image generation. This paper looks at four of the most popular generators, with the aim to find out which one is the best for an inexperienced artist.

@@

	\subsection{Context and background}
	During the year of 2022 there has been a massive increase in the popularity and use of text-to-image generators. "With text-to-image models, people provide a text description and the models produce images matching the description as closely as possible" (\cite{whatIsTextToImageGen}). You will likely have heard of at least one of these if you frequent the internet. This explosion in popularity has led to many different text-to-image applications being made, which on the one hand has given many more options to artists (people who use a text-to-image generator to create and refine an image). However, in our opinion it has also made the landscape more difficult to navigate because of its abundance of choice. This is illustrated by this article which shows that you would have to go through all of the models to find which is the best for each user as this author and most others all base their judgment on their own opinion (\cite{whyUseImageComp}). This paper makes an effort to try and alleviate this problem by picking four of the most used text-to-image generators and comparing them based on their accuracy and process in an objective way. This allows us to help people gain some perspective on which text-to-image generator is best for their needs.

@@

	We expect the best text-to-image generator out of the four we chose to be stable diffusion. This is due to the fact that stable diffusion offers its users more tools and techniques like inpainting (allows you to regenerate a specific area of an image without regenerating the whole image), this allows the user more freedom to fine-tune their result. Given enough time this will yield a more accurate result.

@@

	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{OG}
		\caption{The image the subjects tried to replicate}
		\label{fig:og}
	\end{figure}

	\subsubsection{Scoring generated images}
	With those sets of data, we then plug the images (each iteration) into the image similarity checking API that can give you a numerical value for how close one image is to another (\cite{imageSimilarity}). This way we can see how close each iteration is to the image we are trying to recreate. The number of iterations each application is able to create then gives us insight into how fast the application is and how many iterations you have to do to achieve a certain result. We also gain an idea of how easy to use each application is through this process.

>>>>

	\begin{abstract}
		Following the increase in popularity of artificial intelligence during the year of 2022, there are now many text-to-image generation applications out there. This is great for the field of artificial intelligence as a whole, however it has made knowing which application is best quite hard for artists who are new to text-to-image generation. This paper looks at four of the most popular generators, with the aim to find out which one is the best for an inexperienced artist.

@@

	\subsection{Context and background}
	During the year of 2022 there has been a massive increase in the popularity and use of text-to-image generators. According to \citeauthor{whatIsTextToImageGen} (\citeyear{whatIsTextToImageGen}), text-to-image generators allow people to provide a text description which the generator then uses to produce images matching the description as closely as possible. This increase in popularity has led to many different text-to-image applications being made, which on the one hand has given many more options to artists (people who use a text-to-image generator to create and refine an image). However, in our opinion it has also made the landscape more difficult to navigate because of its abundance of choice. This is illustrated by this article which shows that you would have to go through all of the models to find which is the best for each user as this author and most others all base their judgment on their own opinion (\cite{whyUseImageComp}). This paper makes an effort to try and alleviate this problem by picking four of the most used text-to-image generators and comparing them based on their accuracy and process in an objective way. This allows us to help people gain some perspective on which text-to-image generator is best for their needs.

@@

	We expect the best text-to-image generator out of the four we chose to be stable diffusion. This is due to the fact that stable diffusion offers its users more tools and techniques like inpainting. According to \citeauthor{inpaintingExplaination} (\citeyear{inpaintingExplaination}), inpainting is an AI technique which allows a user to regenerate a missing part of an image. This allows the user more freedom to fine-tune their result. Given enough time this will yield a more accurate result.

@@

	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{OG}
		\caption{The image the subjects tried to replicate (\cite{targetImage})}
		\label{fig:og}
	\end{figure}

	\subsubsection{Scoring generated images}
	We then run the set of generated images, together with the target image through an image similarity checking API that can give you a numerical value for how close one image is to another (\cite{imageSimilarity}). This way we can see how close each iteration is to the image we are trying to recreate. It is able to score the images based on a color histogram and colour location, this is supposed to be resistant to transformations such as scaling and changes in lighting (\cite{imageSimilarityPaper}). The number of iterations each application is able to create then gives us insight into how fast the application is and how many iterations you have to do to achieve a certain result. We also gain an idea of how easy to use each application is through this process.
